Blum axioms and nondeterministic computation of functions

### 2015-12-15 paper.tex
paper.tex: Add the deterministic space and time examples.

\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

@@

>>>>

\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}

@@

\begin{example}
    Given the Gödel numbering $\pi$ of the previous chapter,
    the number of steps a Turing machine needs before halting
    is a complexity measure that satisfies Blum axioms.
    The corresponding function $\Phi$ is
    \begin{equation*}
        \Phi(w, x) = \begin{cases}
            k, &\text{if $w$ takes exactly $k$ steps on $x$ before halting} \\
            \text{undefined}, &\text{if $w$ never halts on $x$}
        \end{cases}
    \end{equation*}
    Then, $\Phi(w, x)$ is defined exactly when $\pi_w(x)$ exists
    (that is, when $w$ halts on $x$),
    and we can determine whether $\Phi(w, x) = k$ for a given $k$
    by simulating $w$ on $x$ for exactly $k$ steps.
\end{example}

\begin{example}
    Space complexity requires some care,
    because a machine may never halt even though it remains confined
    in a limited number of cells.
    Thus, we must leave the space complexity of a machine that never halts undefined.
    Formally, define the function $\Phi$ by
    \begin{equation*}
        \Phi(w, x) = \begin{cases}
            k, &\text{if $w$ uses exactly $k$ tape cells on $x$ before halting} \\
            \text{undefined}, &\text{if $w$ never halts on $x$}
        \end{cases}
    \end{equation*}
    This makes $\Phi$ satisfy the first axiom.
    To satisfy the second axiom,
    we must note that,
    if we limit the number of tape cells to $k$,
    there is a finite number of different computation states
    the machine can be.
    Therefore,
    to determine whether $\Phi(w, x) = k$,
    simulate $w$ on $x$ and keep the whole history of computation.
    If, in any point of the simulation,
    the machine $w$ uses more than $k$ tape cells,
    reject --- the answer is ``no''.
    If any two computation states repeat,
    then the machine will loop and never halt
    --- the answer is ``no'' again.
    Otherwise, the machine must halt
    (since there is a finite number of computation states).
    Now, when the machine halts,
    we just need to make sure it actually used exactly $k$ cells in one of its tapes.
\end{example}
